A Wait-free Queue with Polylogarithmic Step Complexity

### 2023-01-08 podc_paper/correctness-2.tex
working on correctness of GC version

\subsection{Correctness}
Unfortunately, there are enough changes to the algorithm that a new proof of correctness
is required.  Its structure mirrors the proof of the original algorithm, but requires additional
reasoning to ensure GC does not interfere with other routines.

\begin{observation}\label{RBTinsertions}
When a block is inserted into an RBT $T$, its \fld{index} is one greater than the maximum \fld{index} in $T$.
\end{observation}
\begin{proof}
If line \ref{appendLeafGC} inserts a block into a leaf's RBT, then it was created at 
line \ref{enqNewGC} or \ref{deqNewGC} with \fld{index} $h+1$, where $h$ is the maximum \fld{index}
in the leaf's RBT.  When the new block is inserted, the maximum \fld{index} in the RBT is still $h$,
since only the owner of the leaf modifies the leaf's RBT.

If line \ref{refreshInsertGC} inserts a block  into an internal node's RBT $T$,
it is created by the call to \op{CreateBlock} at line \ref{invokeCreateBlockGC} and 
has \fld{index} $h$, which is set at line \ref{refreshReadMax} to  the maximum \fld{index} in $T$ plus one.
\end{proof}

\begin{observation}\label{lastinroot}
Each entry in \var{last} is the \var{index} of a block that has appeared in the root's RBT.
\end{observation}
\begin{proof}
Only line \ref{setlast1} and \ref{setlast2} in \op{FindResponse} write values into \var{last}.
The value written is either $b$ or $b_e \leq b$.
By the precondition of \op{FindResponse}, $\var{root}.\fld{blocks}[b]$ exists, so the index $b$ has appeared
in the root's RBT.
By \Cref{RBTinsertions}, all values less than $b$ have appeared in the root's RBT too.
\end{proof}

\begin{observation}
\label{edirGC} 
If $B$ is a block in the RBT of an internal node $v$, then for $\var{dir} \in \{\fld{left},\fld{right}\}$,
$B.\edir$ was read from the \fld{index} field of some node in the RBT of $v.\var{dir}$.
\end{observation}
\begin{proof}
The \eleft\ and \eright\ fields are set by reading the \fld{index} field of a node in the appropriate child
on line \ref{createEndLeftGC} and \ref{createEndRightGC}.
\end{proof}

\begin{invariant}\label{nonemptyConsec}
For each RBT $T$, 
$T$ is non-empty and the blocks of $T$ have consecutive \fld{index} fields.
\end{invariant}
\begin{proof}
The claim is true initially, since each RBT initially contains only the empty block with \fld{index} 0.
We show that each time a new RBT is constructed, it preserves the invariant.
When a new RBT is created by inserting an element into an existing RBT, \Cref{RBTinsertions}
implies that the invariant is preserved.

Suppose a new RBT is created by a \op{Split}($T,t$) operation $S$ on at line \ref{split} of a \op{Discard}($v,t$), 
and that the invariant held prior to $S$.
Clearly, $S$ preserves the property that indices are consecutive.  To show that the new tree
is non-empty we first prove that the call to \op{Discard}($v,t$) has a parameter $t$ that is at least as big as a value previously
read from the index field of some block in $v$'s RBT.
If \op{Discard} was called from line \ref{discard}, then $t+1$ was read from an entry of \var{last}, so the claim follows from \Cref{lastinroot}.
If \op{Discard} was called from line \ref{discardrecurse1} or \ref{discardrecurse2}, the claim follows
from \Cref{edirGC}.
Since the invariant held prior to $S$, no split before $S$ emptied $v$'s tree.  Therefore,
some block with index at least $t$ is still in $T$ when $S$ occurs.  That block will be in the new tree created by $S$.
\end{proof}

Since RBTs are always non-empty, calls to \op{MaxBlock} are well-defined.  
Throughout the proof, we use $v.\blocks[b]$ to refer to the block with \fld{index} $b$ that appeared
in $v$'s tree at some time during the execution.
It follows from \Cref{RBTinsertions} and \Cref{nonemptyConsec} that this block is unique, if 
it exists.
We also use this notation in the code to indicate that a search of the RBT should be performed for the block
with \fld{index} $b$.
We describe later what should happen if the block is not found by the search because it has been removed
by a GC phase.
\here{remember to come back to this point}

\begin{corollary}\label{maxNonDecreasing}
The maximum \fld{index} in $v$'s RBT is non-decreasing over time.
\end{corollary}
\begin{proof}
The CAS steps (on line \ref{casGC} and \ref{splitCASGC}) that update $v.\fld{blocks}$ always replace 
a tree $T$ by a tree that is built from $T$ by inserting an element and/or removing the elements with 
the smallest \fld{index} values, without emptying the tree.  Thus, the maximum \fld{index} in the tree cannot decrease.
\end{proof}

We now prove the analogue of Lemma \ref{lem::headProgress}.
\begin{lemma}
 If $b>0$ and a block with index $b$ has been inserted into $v.\fld{blocks}$ then 
 $v.\fld{blocks}[b-1].\fld{end\sub{left}} \leq v.\fld{blocks}[b].\fld{end\sub{left}}$ and 
 $v.\fld{blocks}[b-1].\fld{end\sub{right}} \leq v.\fld{blocks}[b].\fld{end\sub{right}}$.
\end{lemma}
\begin{proof}
The block $B$ with index $b$ was installed into $v$'s RBT by the CAS at line \ref{casGC} or \ref{splitCASGC} 
that changed the tree from $T$ to $T_{new}$.
Before this CAS, the tree $T$ was read from $v.blocks$ at line \ref{readTGC} or \ref{readT2GC}, respectively,
before the \op{RefreshUpdate} on line \ref{RU1} or \ref{RU2} that inserted $B$ to get $T_{new}$.
Since the previous block $B'=v.\fld{blocks}[b-1]$ was already in $T$ (by \Cref{RBTinsertions}),
the \op{CreateBlock}$(v,b-1)$ that created $B'$ terminated before the \op{CreateBlock}$(v,b)$ that created $B$ started.
It follows from \Cref{maxNonDecreasing} that the value that line \ref{createEndLeftGC} of \op{CreateBlock}$(v,b-1)$ stores in $B'.\eleft$ is less than or equal to the value that line \ref{createEndLeftGC} of \op{CreateBlock}$(v,b)$ stores in $B.\eleft$.  Similarly, the values stored in $B'.\eright$ and $B.\eright$ at line \ref{createEndRightGC} satisfy the claim.
\end{proof}

\Cref{lem::subblocksDistinct}, \Cref{lem::noDuplicates}, \Cref{lem::sum}, \Cref{blockNotEmpty} 
hold true for the modified algorithm.  Their proofs are identical to those given in 
Section \ref{sec::basicProperties} since they only depend on \Cref{lem::headProgress} 
and the definition of subblocks given in (\ref{defsubblock}).
The following pair of lemmas are analogous to \Cref{successfulRefresh} for the modified \op{Refresh} routine.
We say that an operation has been \emph{propagated} to node $v$ in configuration $C$ if
some block containing the operation has been in $v.blocks$ at or before $C$.

\begin{lemma}\label{nullRefreshGC}
Suppose $v.\fld{blocks}=T$ at some configuration $C$ and \op{RefreshUpdate}($T,v$) is invoked after $C$ and
returns \nl.  Then, all operations that were propagated to $v$'s
children before $C$ have been propagated to $v$ before $C$.
\end{lemma}
\begin{proof}
Suppose an operation $op$ has been propagated to $v$'s child (without loss of generality, $v.\fld{left}$) before $C$.
Then, for some $i$, a block $B$ with \fld{index} $i$ has been in $v.\fld{left.blocks}$ before $C$.
By Lemma \ref{maxNonDecreasing}, 
the maximum \fld{index} in $v.\fld{left.blocks}$ is at least $i$ at all times during the \op{RefreshUpdate}.
So, the value of $\var{new}.\eleft$ chosen at line \ref{createEndLeftGC} of the \op{RefreshUpdate}'s call
to \op{CreateBlock} is at least $i$.

Let $h$ be the value set at line \ref{refreshReadMax} of the \op{RefreshUpdate}.
The \op{RefreshUpdate} returns \nl\ at line \ref{addOPGC}
because the \op{CreateBlock}$(v,h)$  at line \ref{CBnullGC} returns \nl.
By reasoning identical to the last paragraph of \Cref{successfulRefresh},
it follows from the fact that $\var{num\sub{enq}}+\var{num\sub{deq}}=0$ that
the blocks $v.\fld{left.blocks}[v.\fld{blocks}[h-1].\eleft+1..new.\eleft]$ and
$v.\fld{right.blocks}[v.\fld{blocks}[h-1].\eright+1..new.\eright]$
contain no operations.
Since each block contains at least one operation by \Cref{blockNotEmpty}
these ranges must be empty, so $v.\fld{blocks}[h-1].\eleft \geq \var{new}.\eleft\geq i$.
This implies that the block $B$ containing $op$ is a subblock of some block that has appeared in $v.\fld{blocks}$, so $op$ has been propagated to $v$.
\end{proof}

\begin{lemma}\label{successfulRefreshGC}
If a \op{CAS} at line \ref{casGC} or \ref{splitCASGC} modifies $v.\fld{blocks}$ from $T$ to $T'$,
then all operations that were propagated to $v$'s children before $v.\fld{blocks}$ was set to $T$
have been propagated to $v$ in the configuration after the CAS.
\end{lemma}
\begin{proof}
Suppose an operation $op$ has been propagated to $v$'s child (without loss of generality, $v.\fld{left}$) before $v.\fld{blocks}$ was set to $T$.
Then, for some $i$, a block $B$ with \fld{index} $i$ that contains $op$
has been in $v.\fld{left.blocks}$ before $v.\fld{blocks}$ was set to $T$.
Let $c$ be the \op{CAS} instruction that changes $v.\fld{blocks}$ from $T$ to $T'$.
If line \ref{casGC} performs $c$, then the \op{RefreshUpdate} $R$ at line \ref{RU1} that creates
$T'$ began after $T$ was read from $v.\fld{blocks}$ at line \ref{readTGC}.
Otherwise, line \ref{splitCASGC} performs $c$ and the \op{RefreshUpdate} $R$ at line \ref{RU2},
which is invoked after $T$ was read from $v.\fld{blocks}$ at line \ref{readT2GC},
either returned \nl\ or $T'$.

If the \op{RefreshUpdate} $R$ returned \nl, the lemma follows from \Cref{nullRefreshGC}.
So, assume \op{RefreshUpdate} $R$ returned $T'$.
Then $T'$ contains a newly inserted block created by the call
to \op{CreateBlock} on line \ref{invokeCreateBlockGC}.
Since $R$ was invoked after $T$ was stored in $v.\fld{blocks}$, 
the maximum \fld{index} in $v.\fld{left.blocks}$ is at least $i$ throughout $R$, by \Cref{maxNonDecreasing}.
Thus. the \eleft\ field of the new block added to $T'$ by $R$'s call to \op{CreateBlock} is
set to a value greater than or equal to $i$ at line \ref{createEndLeftGC}.
By the definition of subblocks, it follows that the block $B$ containing $op$ is a subblock of a block
that has been in $v.blocks$, as required.
\end{proof}

>>>>

\subsection{Correctness}
Unfortunately, there are enough changes to the algorithm that a new proof of correctness
is required.  Its structure mirrors the proof of the original algorithm, but requires additional
reasoning to ensure GC does not interfere with other routines.
The following observation describes
how the set of blocks in the RBT can be modified.

\begin{lemma}\label{RBTupdates}
Suppose a step of the algorithm changes $v.\fld{blocks}$ from a non-empty tree $T$ to $T'$.
If the set of \fld{index} values  in $T$ is $I$, then the set of \fld{index} values in $T'$ is 
$(I \cap [m-1,\infty))\cup \{\max(I) + 1\}$ for some $m$.
\end{lemma}
\begin{proof}
The RBT of a node is updated only at line \ref{appendLeafGC} or \ref{casGC}.

If line \ref{appendLeafGC} of an \op{Append} operation modifies $v.\fld{blocks}$,
then $v$ is a leaf node, and no other process ever modifies $v.\fld{blocks}$.
$T'$ was obtained from $T$ by calling \op{AddBlock}($T,B$).
$B$ was created either by the \op{Enqueue} or \op{Dequeue} that called \op{Append}.
Either way, $B.\fld{index} = \max(I)+1$.
The \op{AddBlock} that creates $T'$ may optionally \op{Split} the RBT and then add $B$ to it.
So the claim is satisfied.

If line \ref{casGC} of a \op{Refresh} modifies $v.\fld{blocks}$, then $v$ is an internal node.
After reading $T$ from $v.\fld{blocks}$ at line \ref{refreshReadTGC},
then creates the  block \var{new},
and then calls \op{AddBlock}($T,new$) to create $T'$.
Line \ref{createIndexGC} sets $\var{new}.\var{index} = \max(I)+1$.
The \op{AddBlock} that creates $T'$ may optionally \op{Split} the RBT and then add $new$ to it.
So the claim is satisfied.
\end{proof}

Since each RBT starts with a single block with \fld{index} 0, the following is an easy consequence of \Cref{RBTupdates}.
\begin{corollary}\label{nonEmptyIncreasing}
The RBT stored in each node $v$ is never empty and always stores a set of blocks with consecutive indices.
Moreover, its maximum \fld{index} can only increase over time.
\end{corollary}

Since RBTs are always non-empty, calls to \op{MaxBlock} have well-defined answers.  
Throughout the proof, we use $v.\blocks[b]$ to refer to the block with \fld{index} $b$ that appeared
in $v$'s tree at some time during the execution.
It follows from \Cref{RBTupdates} and \Cref{nonemptyConsec} that 
each time a new block appears in $v$'s RBT, its \fld{index} is greater than any block
that has appeared in $v$'s RBT earlier.
Thus, $v.\blocks[b]$ is unique, if it exists.
We also use this notation in the code to indicate that a search of the RBT should be performed for the block
with \fld{index} $b$.
We describe later what should happen if that block is not found by the search because it has been removed
by a GC phase.
\here{remember to come back to this point}

We now establish that definition (\ref{defsubblock}) of a block's subblocks
still makes sense.

%\here{If next obs is only used once, can fold it into the proof that needs it}
%\begin{observation}
%\label{edirGC} 
%If $B$ is a block in the RBT of an internal node $v$, then for $\var{dir} \in \{\fld{left},\fld{right}\}$,
%$B.\edir$ was read from the \fld{index} field of some node in the RBT of $v.\var{dir}$.
%\end{observation}
%\begin{proof}
%The \eleft\ and \eright\ fields are set by reading the \fld{index} field of a node in the appropriate child
%on line \ref{createEndLeftGC} and \ref{createEndRightGC}.
%\end{proof}

We next prove the analogue of Lemma \ref{lem::headProgress}.
\begin{customlemma}{$4'$}\label{lem::headProgressGC}
 If $h>0$ and a block with index $h$ has been inserted into $v.\fld{blocks}$ then 
 $v.\fld{blocks}[h-1].\fld{end\sub{left}} \leq v.\fld{blocks}[h].\fld{end\sub{left}}$ and 
 $v.\fld{blocks}[h-1].\fld{end\sub{right}} \leq v.\fld{blocks}[h].\fld{end\sub{right}}$.
\end{customlemma}
\begin{proof}
The block $B$ with index $h$ was installed into $v$'s RBT by the CAS at line \ref{casGC} 
that changed the tree from $T$ to $T'$.
Before this CAS, line \ref{refreshReadTGC} read the tree $T$ from $v.blocks$,
line \ref{refreshReadMax} found a block $B'$  with \fld{index} $h-1$ in $T$,
and then line \ref{invokeCreateBlockGC} created the block $B$ with $\fld{index}=h$.
Since $B'$ was already in $T$ before $B$ was created, the
\op{CreateBlock}$(v,b-1)$ that created $B'$ terminated before the 
\op{CreateBlock}$(v,b)$ that created $B$ started.
By \Cref{nonEmptyIncreasing}, the value that line \ref{createEndLeftGC} of \op{CreateBlock}$(v,b-1)$ 
stores in $B'.\eleft$ is less than or equal to the value that line \ref{createEndLeftGC} 
of \op{CreateBlock}$(v,b)$ stores in $B.\eleft$.  
Similarly, the values stored in $B'.\eright$ and $B.\eright$ at line \ref{createEndRightGC} satisfy the claim.
\end{proof}

\Cref{lem::subblocksDistinct}, \Cref{lem::noDuplicates}, \Cref{lem::sum}, \Cref{blockNotEmpty} 
hold for the modified algorithm.  Their proofs are identical to those given in 
Section \ref{sec::basicProperties} since they  depend only on \Cref{lem::headProgress} (which can be replaced by \Cref{lem::headProgressGC})
and the definition of subblocks given in (\ref{defsubblock}).
Next, we prove an analogue of \Cref{successfulRefresh}.
We say a node $v$ \emph{contains} an operation
if some block containing the operation has previously appeared in $v$'s RBT (even if 
the block has been removed from the RBT by a subsequent \op{Split}).

\begin{customlemma}{$9'$}
\label{successfulRefreshGC}
Let $R$ be a call to \op{Refresh}($v$) that performs a successful \op{CAS} on line \ref{casGC} (or terminates at line \ref{addOPGC}).
In the configuration after that CAS (or termination, respectively), $v$ contains all operations that $v$'s children contained 
when $R$ executed line~\ref{refreshReadTGC}.
\end{customlemma}
\begin{proof}
Suppose $v$'s child (without loss of generality, $v.\fld{left}$) contained an operation $op$ 
when $R$ executed line \ref{refreshReadTGC}.
Let $i$ be the index of the block containing $op$ that was in $v$'s RBT before 
$R$ executed line \ref{refreshReadTGC}.
We consider two cases.

Suppose $R$'s call to \op{CreateBlock} returns a new block $B'$.  Then, $R$'s \op{CAS} at line \ref{cas} 
installs $B'$ in $v.\fld{blocks}$.
The \op{CreateBlock} set $B'.\eleft$ to the maximum \fld{index} in $v.\fld{left}$'s RBT at line \ref{createEndLeftGC}.
By \Cref{nonEmptyIncreasing}, this maximum \fld{index} is bigger than $i$.
By the definition of subblocks, some block in $v$ contains $B$ as a subblock
and therefore $v$ contains $op$.

Now suppose $R$'s call to \op{CreateBlock} returns \nl, causing $R$ to terminate at line \ref{addOPGC}.
Let $h$ be the maximum \fld{index} in $T$ plus 1.
By reasoning identical to the last paragraph of \Cref{successfulRefresh}'s proof,
it follows from the fact that $\var{num\sub{enq}}+\var{num\sub{deq}}=0$ at line \ref{testEmptyGC}
that the blocks $v.\fld{left.blocks}[v.\fld{blocks}[h-1].\eleft+1..new.\eleft]$ and
$v.\fld{right.blocks}[v.\fld{blocks}[h-1].\eright+1..new.\eright]$
contain no operations.
By \Cref{blockNotEmpty}, each block contains at least one operation, so
these ranges must be empty, and $v.\fld{blocks}[h-1].\eleft \geq \var{new}.\eleft\geq i$.
This implies that the block $B$ containing $op$ is a subblock of some block that has appeared
in $v$'s RBT, so $op$ is contained in $v$.
\end{proof}

### 2023-01-08 podc_paper/correctness.tex
working on correctness of GC version

Now suppose $v$ is an internal node. By the definition of subblocks in (\ref{defsubblock}), the
subblocks of $v.\fld{blocks}[1..i]$ are $v.\fld{left.blocks}[1..B.\eleft]$ 
and $v.\fld{right.blocks}[1..B.\eright]$.
Thus, the enqueues in $E(v.\fld{blocks}[0])\cdots E(v.\fld{blocks}[i])$ are those in
$E(v.\fld{left.blocks}[0]) \cdots E(v.\fld{left.blocks}[B.\eleft])$ and
$E(v.\fld{left.blocks}[0]) \cdots E(v.\fld{left.blocks}[B.\eright])$.
By the hypothesis, the total number of these enqueues is $v.\fld{left.blocks}[B.\eleft].\fld{sum\sub{enq}} + v.\fld{right.blocks}[B.\eright].\fld{sum\sub{enq}}$, which is the value that line \ref{createSumEnq} stored in $B.\fld{sum\sub{enq}}$ when $B$ was created.
The proof for \fld{sum\sub{deq}} (stored on line~\ref{createSumDeq}) is similar.
\end{proof}

@@

\begin{lemma}\label{successfulRefresh}
Let $R$ be a call to \op{Refresh}($v$) that performs a successful \op{CAS} on line \ref{cas} (or terminates at line \ref{addOP}).
In the configuration after that CAS (or termination, respectively), $v$ contains all operations that $v$'s children contained 
when $R$ executed line~\ref{readHead}.
\end{lemma}
\begin{proof}
Suppose $v$'s child (without loss of generality, $v.\fld{left}$) contained an operation $op$ 
in the configuration $C$ immediately before $R$ executed line \ref{readHead}.
Then for some $i$, the block $B$ in $v.\fld{left.blocks}[i]$ contains $op$.
By \Cref{nonDecreasingHead} and \Cref{lem::headProgress}, the value of $childHead$ that $R$ reads from
$v.\fld{left.head}$ in line \ref{readChildHead} is at least $i$.
If it is equal to $i$, $R$ calls \op{Advance} at line \ref{helpAdvance}, which ensures that 
$v.\fld{left.head} > i$.
Then, $R$ calls \op{CreateBlock}($v,h$) in line \ref{invokeCreateBlock}, where $h$ is the value $R$ reads at line \ref{readHead}.
\op{CreateBlock} reads a value greater than $i$ from $v.\fld{left.head}$ at line \ref{createEndLeft}.
Thus, $new.\eleft \geq i$.  We consider two cases.

Suppose $R$'s call to \op{CreateBlock} returns the new block $B'$ and $R$'s \op{CAS} at line \ref{cas} 
installs $B'$ in $v.\fld{blocks}$.
Then, $B$ is a subblock of some block in $v$, since  $B'.\eleft$ is greater than or equal to $B$'s index
in $v.\fld{left.blocks}$.
Hence $B'$ contains $op$, as required.

>>>>

Now suppose $v$ is an internal node. By the definition of subblocks in (\ref{defsubblock}) and \Cref{lem::headProgress}, the
subblocks of $v.\fld{blocks}[1..i]$ are $v.\fld{left.blocks}[1..B.\eleft]$ 
and $v.\fld{right.blocks}[1..B.\eright]$.
Thus, the enqueues in $E(v.\fld{blocks}[0])\cdots E(v.\fld{blocks}[i])$ are those in
$E(v.\fld{left.blocks}[0]) \cdots E(v.\fld{left.blocks}[B.\eleft])$ and
$E(v.\fld{left.blocks}[0]) \cdots E(v.\fld{left.blocks}[B.\eright])$.
By the hypothesis, the total number of these enqueues is $v.\fld{left.blocks}[B.\eleft].\fld{sum\sub{enq}} + v.\fld{right.blocks}[B.\eright].\fld{sum\sub{enq}}$, which is the value that line \ref{createSumEnq} stored in $B.\fld{sum\sub{enq}}$ when $B$ was created.
The proof for \fld{sum\sub{deq}} (stored on line~\ref{createSumDeq}) is similar.
\end{proof}

@@

\begin{lemma}\label{successfulRefresh}
Let $R$ be a call to \op{Refresh}($v$) that performs a successful \op{CAS} on line \ref{cas} (or terminates at line \ref{addOP}).
In the configuration after that CAS (or termination, respectively), $v$ contains all operations that $v$'s children contained 
when $R$ executed line~\ref{readHead}.
\end{lemma}
\begin{proof}
Suppose $v$'s child (without loss of generality, $v.\fld{left}$) contained an operation $op$ 
when $R$ executed line \ref{readHead}.
Let $i$ be the index such that the block $B=v.\fld{left.blocks}[i]$ contains $op$.
By \Cref{nonDecreasingHead} and \Cref{lem::headProgress}, the value of $childHead$ that $R$ reads from
$v.\fld{left.head}$ in line \ref{readChildHead} is at least $i$.
If it is equal to $i$, $R$ calls \op{Advance} at line \ref{helpAdvance}, which ensures that 
$v.\fld{left.head} > i$.
Then, $R$ calls \op{CreateBlock}($v,h$) in line \ref{invokeCreateBlock}, where $h$ is the value $R$ reads at line \ref{readHead}.
\op{CreateBlock} reads a value greater than $i$ from $v.\fld{left.head}$ at line \ref{createEndLeft}.
Thus, $new.\eleft \geq i$.  We consider two cases.

Suppose $R$'s call to \op{CreateBlock} returns the new block $B'$ and $R$'s \op{CAS} at line \ref{cas} 
installs $B'$ in $v.\fld{blocks}$.
Then, $B$ is a subblock of some block in $v$, since  $B'.\eleft$ is greater than or equal to $B$'s index
in $v.\fld{left.blocks}$.
Hence $v$ contains $op$, as required.
